### 2013-02-20 writeup/summary.tex
compressed summary table

\begin{figure}[h]
\footnotesize
\begin{tabular}{lccl}
\multicolumn{4}{c}{Number of turkers per language}\\
\hline\hline
&\multicolumn{2}{c}{Avg. turker quality (\# Turkers)}&Primary locations\\
&In region&Out of region&of turkers in region\\
\hline\hline
Hindi&\textbf{0.54} (299)&0.47 (63)&India (285) UAE (5) UK (5) \\
Tamil&\textbf{0.55} (278)&0.54 (26)&India (271) US (3) Canada (2) \\
Malayalam&0.48 (235) &\textbf{0.55} (49)&India (224) UAE (6) US (3) \\
Spanish&\textbf{0.27} (205)&0.27 (39)&US (131) Mexico (18) Spain (14) \\
French&0.29 (183) &\textbf{0.31} (34)&India (68) US (48) France (25) \\
German&\textbf{0.35} (99)&0.31 (71)&Germany (51) US (29) Austria (9) \\
Italian&\textbf{0.32} (98)&0.27 (68)&Italy (44) US (32) Romania (9) \\
Chinese&\textbf{0.42} (123)&0.41 (28)&US (80) Singapore (15) China (14) \\
Telugu&\textbf{0.48} (102)&0.41 (35)&India (98) US (3) UAE (1) \\
Malay&\textbf{0.30} (15)&0.28 (116)&Malaysia (8) Singapore (4) US (2) \\
Turkish&\textbf{0.30} (83)&0.27 (47)&Turkey (39) US (20) Macedonia (12) \\
Kannada&0.64 (106) &\textbf{0.65} (21)&India (106) US (1) \\
Portuguese&0.20 (107) &\textbf{0.24} (19)&Brazil (46) Portugal (32) US (15) \\
Arabic&0.45 (62) &\textbf{0.45} (64)&Egypt (19) Jordan (17) Morocco (9) \\
Romanian&0.29 (87) &\textbf{0.33} (37)&Romania (77) US (6) Moldova (3) \\
Swedish&\textbf{0.30} (58)&0.29 (66)&US (28) Sweden (23) Finland (3) \\
Czech&0.24 (52) &\textbf{0.25} (71)&US (19) Czechrepublic (15) Serbia (5) \\
Irish&\textbf{0.32} (58)&0.26 (64)&US (41) Ireland (13) UK (4) \\
Amharic&0.21 (17) &\textbf{0.42} (105)&US (14) Ethiopia (3) \\
Russian&0.15 (72) &\textbf{0.16} (49)&US (38) Russia (8) Moldova (7) \\
\hline\hline
\end{tabular}
\end{figure}