### 2013-02-20 writeup/summary.tex
new quality scores

\begin{figure}[h]
\footnotesize
\begin{tabular}{llll}
\multicolumn{4}{c}{Number of turkers per language}\\
\hline\hline
&\multicolumn{2}{c}{Avg. turker quality (\# Turkers)}&Primary locations\\
&In region&Out of region&of turkers in region\\
\hline\hline
Hindi&\textbf{0.410} (299)&0.406 (63)&India (285) UAE (5) UK (5) \\
Tamil&0.339 (278) &\textbf{0.482} (26)&India (271) US (3) Canada (2) \\
Malayalam&\textbf{0.534} (235)&0.505 (49)&India (224) UAE (6) US (3) \\
Spanish&0.561 (205) &\textbf{0.682} (39)&US (131) Mexico (18) Spain (14) \\
French&\textbf{0.636} (183)&0.541 (34)&India (68) US (48) France (25) \\
German&\textbf{0.691} (99)&0.678 (71)&Germany (51) US (29) Austria (9) \\
Italian&0.650 (98) &\textbf{0.716} (68)&Italy (44) US (32) Romania (9) \\
Chinese&0.165 (123) &\textbf{0.207} (28)&US (80) Singapore (15) China (14) \\
Telugu&\textbf{0.582} (102)&0.582 (35)&India (98) US (3) UAE (1) \\
Malay&0.630 (15) &\textbf{0.649} (116)&Malaysia (8) Singapore (4) US (2) \\
Turkish&0.593 (83) &\textbf{0.634} (47)&Turkey (39) US (20) Macedonia (12) \\
Kannada&0.155 (106) &\textbf{0.284} (21)&India (106) US (1) \\
Portuguese&0.599 (107) &\textbf{0.651} (19)&Brazil (46) Portugal (32) US (15) \\
Arabic&\textbf{0.278} (62)&0.217 (64)&Egypt (19) Jordan (17) Morocco (9) \\
Romanian&0.653 (87) &\textbf{0.668} (37)&Romania (77) US (6) Moldova (3) \\
Swedish&0.497 (58) &\textbf{0.504} (66)&US (28) Sweden (23) Finland (3) \\
Czech&0.536 (52) &\textbf{0.571} (71)&US (19) Czechrepublic (15) Serbia (5) \\
Irish&\textbf{0.645} (58)&0.605 (64)&US (41) Ireland (13) UK (4) \\
Amharic&0.080 (17) &\textbf{0.138} (105)&US (14) Ethiopia (3) \\
Russian&0.202 (72) &\textbf{0.275} (49)&US (38) Russia (8) Moldova (7) \\
\hline\hline
\end{tabular}
\end{figure}